---
title: Holiday Movies & Their Changes
author: Yellow Echidna <br> Nayeon Kwon, Aishwarya Gupta, Yuxuan Chen
date: 2/29/2024
format: revealjs
execute:
  fig-width: 7
  fig-asp: 0.618
  fig-retina: 2
  dpi: 150
  fig-align: center
  out-width: 80%
knitr:
  opts_chunk:
    fig.showtext: true
---

## Introduction

Dataset we used: holiday_movies.csv

::: {.nonincremental}
- identifiers
- titles
- release years
- runtime
- genres
- ratings
- vote counts
- boolean flags for specific holiday keywords in the title ("Christmas," "Hanukkah," "Kwanzaa," and "holiday")
:::

# Question 1

## Introduction

What's the average rating and counts of different holiday movie genres across the decades, and which genres are the most popular for each decade?

## Approach
::: {.nonincremental}
- Current Variables: 'average_rating', 'genres', and 'year'
- New Variables: decades (grouped from 'year'), average rating, number of movies (from 'genres')

...

- Stacked bar charts --> Visualize genre distribution over the decades
- Line chart --> Visualize average ratings over decades
- Faceted charts --> Visualize individual genres
:::

## Analysis

::: {.panel-tabset}
### Stacked bar chart
```{r}
#| fig-width: 20
#| fig-height: 75

library(tidyverse)
library(lubridate)
library(ggplot2)
library(dplyr)
library(caret)
library(broom)
library(plotly)

url <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-12-12/holiday_movies.csv"

holiday_movies <- read.csv(url(url))

holiday_movies_q1 <- holiday_movies |>
  mutate(
    decade = 10 * (year %/% 10),  
    genre_list = strsplit(as.character(genres), ",\\s*")
  )


# Count the number of movies per genre for each decade
genre_count_decade <- holiday_movies_q1 |>
  unnest(genre_list) |>
  group_by(decade, genre_list) |>
  summarise(count = n())

# Calculate the average rating for each genre within each decade
genre_rating_decade <- holiday_movies_q1 |>
  unnest(genre_list) |>
  group_by(decade, genre_list) |>
  summarise(average_rating = mean(average_rating, na.rm = TRUE))

# Visualize the genre distribution over the decades with a stacked bar chart
ggplot(genre_count_decade, aes(x = as.factor(decade), y = count, fill = genre_list)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(
    title = "Genre distribution for each decade",
    x = "Decade", y = "Number of movies", 
    fill = "Genre"
    ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.title.position = "plot"
    )
```

### Facet Bar
```{r}
#| fig-width: 20
#| fig-height: 75

# Facet wrap for each genre
ggplot(genre_count_decade, aes(x = as.factor(decade), y = count, fill = genre_list)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  theme_minimal() +
  facet_wrap(facets = vars(genre_list)) +
  scale_y_continuous() +
  labs(
    title = "Genre distribution for each decade",
    x = "Decade", y = "Number of movies", 
    fill = "Genre"
    ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 5),
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.title.position = "plot"
    )
```

### Line Chart
```{r}
#| fig-width: 20
#| fig-height: 75

# Visualize average ratings over decades with a line chart
ggplot(genre_rating_decade, aes(x = as.factor(decade), y = average_rating, group = genre_list, color = genre_list)) +
  geom_line() +
  theme_minimal() +
  labs(
    title = "Average rating change for each decade",
    x = "Decade", y = "Average Rating",
    color = "Genre"
    ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.title.position = "plot"
        )
```


### Facet Line
```{r}
#| fig-width: 20
#| fig-height: 75

# Facet wrap each average rating line
ggplot(genre_rating_decade, aes(x = as.factor(decade), y = average_rating, group = genre_list, color = genre_list)) +
  geom_line() +
  theme_minimal() +
  facet_wrap(facets = vars(genre_list)) +
  scale_y_continuous() +
  labs(
    title = "Average rating change for each genre",
    x = "Decade", y = "Average Rating",
    color = "Genre"
    ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 5),
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.title.position = "plot",
    legend.position = ""
        )
```
:::

## Discussion

:::: {.columns}

::: {.column width="20%"}
Count

Comedy
Drama
Comedy
Drama
Drama
Family
Family
Family
Comedy
Comedy
Romance
:::

::: {.column width="20%"}
Decades 

1920s
1930s
1940s
1950s
1960s
1970s
1980s
1990s
2000s
2010s
2020s
:::

::: {.column width="20%"}
Avg Rating

Comedy
Family
Music
Family
Animation
Music
History
War   
War
Biography
Documentary
:::

::::

# Question 2

## Introduction

How did Christmas, Hanukkah, and Kwanzaa movie distribution change over the years? How different are their average ratings?

## Approach

::: {.nonincremental}
- Variables: 'christmas', 'hanukkah', 'kwanzaa', 'year', and 'average_rating'.

...

- Line chart --> Visualize each holiday's movie change over years / compare trends easily
- Bar chart --> Visualize average rating per holiday / straightforward comparison
:::

## Analysis

```{r}
#| fig-width: 20
#| fig-height: 30

library(tidyverse)
library(ggplot2)
library(gridExtra)

holiday_movies_line <- holiday_movies %>%
  mutate(across(c(christmas, hanukkah, kwanzaa), as.numeric)) %>%
  group_by(year) %>%
  summarise(
    christmas_movies = sum(christmas),
    hanukkah_movies = sum(hanukkah),
    kwanzaa_movies = sum(kwanzaa)
  ) %>%
  pivot_longer(cols = ends_with("_movies"), names_to = "holiday", values_to = "num_movies", names_prefix = "")

line_plot <- ggplot(holiday_movies_line, aes(x = year, y = num_movies, color = holiday)) +
  geom_line() +
  labs(
    title = "Holiday Movie Distribution Over Years",
    x = "Year", y = "Number of Movies", 
    color = "Holiday"
    ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.title.position = "plot",
    legend.position = "bottom"
  )

holiday_movies_bar <- holiday_movies %>%
  pivot_longer(cols = c(christmas, hanukkah, kwanzaa), names_to = "holiday_type", values_to = "is_holiday") %>%
  filter(is_holiday) %>%
  group_by(holiday_type) %>%
  summarise(average_rating = mean(average_rating, na.rm = TRUE))

bar_plot <- ggplot(holiday_movies_bar, aes(x = holiday_type, y = average_rating, fill = holiday_type)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Average Ratings",
    x = "Holiday", y = "Average Rating", 
    fill = "Holiday") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = rel(0.9)),
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.title.position = "plot",
    legend.position = ""
  )

grid.arrange(line_plot, bar_plot, ncol = 2, widths = c(5, 2))
```

## Discussion - Line plot

Holiday Movie Distribution Over Years

::: {.nonincremental}
- Significant increase in Christmas movies & sharp rise in recent years
- Relative scarcity for Hanukkah and Kwanzaa movies throughout the years
- Stronger cultural or commercial emphasis on Christmas-themed content in the film industry
:::

## Discussion - Bar plot

Average Ratings

::: {.nonincremental}
- Kwanzaa --> Hanukkah --> Christmas movies
- Potential quality-over-quantity scenario (lesser-produced holiday movies receive more favorable reviews)
- Potential of reflecting a niche audience's rating behavior
:::

## Q&A

Thank you for listening! Any questions?
